---
title: " select"
author:
- familyname: Shi
  othernames: Jingyi
  address: Monash University
  email: jshi0067@student.monash.edu
  correspondingauthor: true
  qualifications:  32267886
- familyname: Sun
  othernames: Miao
  address: Monash University
  email: msun0008@student.monash.edu
  correspondingauthor: true
  qualifications: 28380584
- familyname: Ou
  othernames: Jingwen
  address: Monash University
  email: jouu0006@student.monash.edu
  correspondingauthor: true
  qualifications:  32269633
- familyname: Qin
  othernames: Yu
  address: Monash University
  email: yqin0016@student.monash.edu
  correspondingauthor: true
  qualifications:  32606745
department: Faculty of Business and Statistic
organization: Monash University ETC5513 Group Really3Q
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    citation_package: biblatex
    toc: false
---

### Benin and Mozambique

#### Question 1: What are top 5 reasons to cause death in Benin and Mozambique, and what may be the reason for it?
#### Question 2: What is the development trend of these reasons in the above question?
```{r,eval=TRUE,echo=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(knitr)
library(gridExtra)
```
#### Findings
```{r, eval=TRUE,echo=FALSE}
annual_number_of_deaths_by_cause <- read_csv("Data/annual-number-of-deaths-by-cause.csv", 
    col_types = cols(`Number of executions` = col_double()))
```
```{r Benin top five death rate, eval=TRUE,echo=FALSE}
disease_data_Benin <- annual_number_of_deaths_by_cause %>% 
  select(1,3,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36) %>% 
  filter(Country %in% c("Benin"),
         (Year >= 2000)) %>% 
  pivot_longer(cols =  -c(1:2), 
               names_to = "Death_cause", 
               values_to = "Count")%>%
  group_by(Death_cause) %>% 
  summarise(Death_amount=sum(Count, na.rm = TRUE))%>%
arrange(desc(Death_amount))%>%
  head(n=5)%>%
  kable(caption = "The top 5 death causes of Benin from 2000 until now")
  
```


```{r Mozambique top five death rate, eval=TRUE,echo=FALSE}
disease_data_Mozambique <- annual_number_of_deaths_by_cause %>% 
  select(1,3,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36) %>% 
  filter(Country %in% c("Mozambique"),
         (Year >= 2000)) %>% 
  pivot_longer(cols =  -c(1:2), 
               names_to = "Death_cause", 
               values_to = "Count")%>%
  group_by(Death_cause) %>% 
  summarise(Death_amount=sum(Count, na.rm = TRUE))%>%
arrange(desc(Death_amount))%>%
  head(disease_data_Mozambique,n=5)%>%
  kable(caption = "The top 5 death causes of Mozambique from 2000 until now")
```
According to that table \@ref(tab: Benin top five death rate), Top five death reasons for Benin from 2000 until now are: Malaria, Neonatal disorders, Lower respiratory infections, cardiovascular disease, Diarrheal diseases.
Meanwhile, according to table \@ref(tab: Mozambique top five death rate) top five death reasons for Mozambique from 2000 unitl now are: HIV/AIDS, Malaria,Neonatal disorders, Cardiovascular diseases,Tuberculosis.
Different from other high-incoem countries with Cardiovascular diseases as the main cause, these two low-income countries's death rate may caused by infectious disease and Neonatal disorders, due to the poor medical system and lack of vaccine and hospital according to @mbaye2019telling.

```{r topfiveplot,fig.cap = "figure of top5 death reason in Benin and Mozambique", eval=TRUE,echo=FALSE}
disease_data_Benin_plot <- annual_number_of_deaths_by_cause %>% 
  select(1,3,9,16,17,18,22) %>% 
  filter(Country %in% c("Benin"),
         (Year >= 2000)) %>% 
  pivot_longer(cols =  -c(1:2), 
               names_to = "Death_cause", 
               values_to = "Count")

disease_data_Mozambique_plot <- annual_number_of_deaths_by_cause %>% 
  select(1,3,9,13,15,16,18) %>% 
  filter(Country %in% c("Mozambique"),
         (Year >= 2000)) %>% 
  pivot_longer(cols =  -c(1:2), 
               names_to = "Death_cause", 
               values_to = "Count")
  
  


P1 <- ggplot(disease_data_Benin_plot, aes(x=Year, 
 y= Count,
 color=Death_cause))+geom_line(size=1)+ggtitle("Trends of top five death rate in Benin")

P2 <- ggplot(disease_data_Mozambique_plot, aes(x=Year, 
 y= Count,
 color=Death_cause))+geom_line(size=1)+ggtitle("Trends of top five death rate in Mozambique")
 
 
gridExtra::grid.arrange(P1,P2)
 
  
  

```


#### Summary
According to that table \@ref(tab: Benin top five death rate), Top five death reasons for Benin from 2000 until now are: Malaria, Neonatal disorders, Lower respiratory infections, cardiovascular disease, Diarrheal diseases.
Meanwhile, according to table \@ref(tab: Mozambique top five death rate) top five death reasons for Mozambique from 2000 unitl now are: HIV/AIDS, Malaria,Neonatal disorders, Cardiovascular diseases,Tuberculosis.
Comparing to other countries, Benin and Mozambique have totally different results. For those high-income or developed countries and  middle-income countries, the main death causes are Cardiovascular disease, cancers, or diabetes mellitus, which are mainly caused by getting old, or unhealthy lifestyle. However, in Benin and Mozambique the main cause of death are different types and infectious diseases and neonatal disorders. The reason for this phenomena is the bad circumstance and  the poor medical system in these countries, also they are lack of vaccines and medicines, according to @mbaye2019telling. Many people could have been saved if they have a good treatment. For example, Malaria  is mainly transmitted by mosquitoes, and in Africa there are so many mosquitoes. However people can be saved if they have specific drugs like artemisinin. In this may, Malaria, a not very severe disease in other countries, but become the top death causes in Benin and Mozambique. 
According to figure graph\@ref(fig:topfiveplot), except Diarrheal disease, all the other four diseases were increasing from 2000 until now.Lower respiratory infections death reached the peak around 2015, and began to decrease in recent years.
For Mozambique, the death rate of Cardiovascular disease was increasing from 2000 until now. Death number for HIV and Tuberculosis reached the peak in around 2013 and began to drop.The death number of Malaria and Neonatal disorder was deceasing from 2000 until now.
In order to help these low-income countries get out of these difficulties, and save more life, high-income developed countries should help them by serving more medicines and vaccines. Also, more medical teams should support Africa and help patients in those low-income countries.
To make these tables and figure, I can also cite R packages as follows @tidyverse, @ggplot2, @gridExtra.

## Conclusion
In conclusion, we can see that in high-income or developed countries like Australia, Switzerland, and Germany, the main death causes are cardiovascular diseases and they are caused by bad lifestyles and a long life time. People need to pay more attention to their mental health,and have a healthy diet. For those middle-income countries like China, India, and Mexico, the main death reason are  cardiovascular and cancer, and the death reason is very related to the GDP of these countries. For those low-income countries, like Benin and Mozambique, people are suffering in many kinds of infectious diseases. We hope more high-income and developed countries can help those poor counties by severing more medical help, medicines and vaccines. All the countries in the world should collaborate with each other and make more people have happiness and healthy life. 