---
title: " select"
author:
- familyname: Shi
  othernames: Jingyi
  address: Monash University
  email: jshi0067@student.monash.edu
  correspondingauthor: true
  qualifications:  32267886
- familyname: Sun
  othernames: Miao
  address: Monash University
  email: msun0008@student.monash.edu
  correspondingauthor: true
  qualifications: 28380584
- familyname: Ou
  othernames: Jingwen
  address: Monash University
  email: jouu0006@student.monash.edu
  correspondingauthor: true
  qualifications:  32269633
- familyname: Qin
  othernames: Yu
  address: Monash University
  email: yqin0016@student.monash.edu
  correspondingauthor: true
  qualifications:  32606745
department: Faculty of Business and Statistic
organization: Monash University ETC5513 Group Really3Q
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    citation_package: biblatex
    toc: false
---

### Benin and Mozambique

#### Question 1: What are top 5 reasons to cause death in Benin and Mozambique, and what may be the reason for it?
#### Question 2: What is the development trend of these reasons in the above question?
```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(knitr)
library(gridExtra)
```
#### Findings
```{r}
annual_number_of_deaths_by_cause <- read_csv("Data/annual-number-of-deaths-by-cause.csv", 
    col_types = cols(`Number of executions` = col_double()))
```
```{r Benin_and_Mozambique_death}
Benin_and_Mozambique <- annual_number_of_deaths_by_cause %>%
  filter(Country %in% c("Benin",
                        "Mozambique"))%>%
  filter(Year >= 2000)%>%
  group_by(Country)%>%
  summarise("Meningitis" = sum(`Meningitis`), 
            "Alzheimer's disease and other dementias" = sum(`Alzheimer's disease and other dementias`),
            "Parkinson's disease" = sum(`Parkinson's disease`),
            "Nutritional deficiencies" = sum(`Nutritional deficiencies`),
            "Malaria" = sum(`Malaria`),
            "Drowning" = sum(`Drowning`),
            "Interpersonal violence" = sum(`Interpersonal violence`),
            "Maternal disorders" = sum(`Maternal disorders`),
            "HIV/AIDS" = sum(`HIV/AIDS`),
            "Drug use disorders" = sum(`Drug use disorders`),
            "Tuberculosis" = sum(`Tuberculosis`),
            "Cardiovascular diseases" = sum(`Cardiovascular diseases`),
            "Lower respiratory infections" = sum(`Lower respiratory infections`),
            "Neonatal disorders" = sum(`Neonatal disorders`),
            "Alcohol use disorders" = sum(`Alcohol use disorders`),
            "Self-harm" = sum(`Self-harm`),
            "Exposure to forces of nature" = sum(`Exposure to forces of nature`),
            "Diarrheal diseases" = sum(`Diarrheal diseases`),
            "Environmental heat and cold exposure" = sum(`Environmental heat and cold exposure`),
            "Neoplasms" = sum(`Neoplasms`),
            "Conflict and terrorism" = sum(`Conflict and terrorism`),
            "Diabetes mellitus" = sum(`Diabetes mellitus`),
            "Chronic kidney disease" = sum(`Chronic kidney disease`),
            "Poisonings" = sum(`Poisonings`),
            "Protein-energy malnutrition" = sum(`Protein-energy malnutrition`),
            "Terrorism (deaths)" = sum(`Terrorism (deaths)`),
            "Road injuries" = sum(`Road injuries`),
            "Chronic respiratory diseases" = sum(`Chronic respiratory diseases`),
            "Cirrhosis and other chronic liver diseases" = sum(`Cirrhosis and other chronic liver diseases`),
            "Digestive diseases" = sum(`Digestive diseases`),
            "Fire, heat, and hot substances" = sum (`Fire, heat, and hot substances`),
            "Acute hepatitis" = sum(`Acute hepatitis`)
            )%>%
  
  print()
```

According to that table \@ref(tab: Benin_and_Mozambique_death), Top five death reasons for Benin from 2000 until now are: Malaria, Neonatal disorders, Lower respiratory infections, cardiovascular disease, Diarrheal diseases.
Meanwhile, top five death reasons for Mozambique from 2000 unitl now are: HIV/AIDS, Malaria,Neonatal disorders, Cardiovascular diseases,Tuberculosis.
Different from other high-incoem countries with Cardiovascular diseases as the main cause, these two low-income countries's death rate may caused by infectious disease and Neonatal disorders, due to the poor medical system and lack of vaccine and hospital according to @mbaye2019telling.

```{r Benintopfiveplot,fig.cap = "figure of top5 death reason in Benin"}
Top_five_Benin <- annual_number_of_deaths_by_cause %>%
 filter(Country == "Benin")%>%
  filter(Year >= 2000)%>%
   select( "Country",
          "Year",
          "Malaria", 
          "Neonatal disorders", 
          "Lower respiratory infections", 
          "Cardiovascular diseases", 
          "Diarrheal diseases"
         )
  

P1 <- ggplot(Top_five_Benin,
             aes(x = Year))+
  geom_line(aes (y=Malaria),
            size=1)
P2 <- ggplot(Top_five_Benin,
             aes(x = Year))+
  geom_line(aes (y= `Neonatal disorders`),
            size=1)
P3 <- ggplot(Top_five_Benin,
             aes(x = Year))+
  geom_line(aes (y= `Lower respiratory infections`),
            size=1)
P4 <- ggplot(Top_five_Benin,
             aes(x = Year))+
  geom_line(aes (y= `Cardiovascular diseases`),
            size=1)
P5 <- ggplot(Top_five_Benin,
             aes(x = Year))+
  geom_line(aes (y= `Diarrheal diseases`),
            size=1)
gridExtra::grid.arrange(P1,P2,P3,P4,P5)

```


```{r Mozambiquetopfiveplot,fig.cap = "figure of top5 death reason in Mozambique"}
Top_five_Mozambique <- annual_number_of_deaths_by_cause %>%
 filter(Country == "Mozambique")%>%
  filter(Year >= 2000)%>%
   select( "Country",
          "Year",
          "HIV/AIDS", 
          "Malaria", 
          "Neonatal disorders", 
          "Cardiovascular diseases", 
          "Tuberculosis"
         )
  

P6 <- ggplot(Top_five_Mozambique,
             aes(x = Year))+
  geom_line(aes (y=`HIV/AIDS`),
            size=1)
P7 <- ggplot(Top_five_Mozambique,
             aes(x = Year))+
  geom_line(aes (y= `Malaria`),
            size=1)
P8 <- ggplot(Top_five_Mozambique,
             aes(x = Year))+
  geom_line(aes (y= `Neonatal disorders`),
            size=1)
P9 <- ggplot(Top_five_Mozambique,
             aes(x = Year))+
  geom_line(aes (y= `Cardiovascular diseases`),
            size=1)
P10 <- ggplot(Top_five_Mozambique,
             aes(x = Year))+
  geom_line(aes (y= `Tuberculosis`),
            size=1)
gridExtra::grid.arrange(P6,P7,P8,P9,P10)
```


#### Summary
According to figure graph\@ref(fig:Benintopfiveplot), accept Diarrheal disease, all the other four disease is increasing from 2000 until now.Lower respiratory infections death reached the peak around 2015, and begin to decrease in recent years.
According to figure graph\@ref(fig:Mozambiquetopfiveplot), For Mozambique, the death rate of Cardiovascular disease was increasing from 2000 unitl now. Death number for HIV and Tuberculosis reached the peak in around 2013 and begin to drop.The death number of Malaria and Neonatal disorder is deceasing from 2000 until now.
In order to help these low-income countries get out of these difficulties, and save more life, high-income developed countries should help them by serving more medicine and vaccine. Also, more medical teams should support Africa and help patients in those low-income countries.
To make these tables and figure, I can also cite R packages as follows @tidyverse, @ggplot2, @gridExtra.